Notes

### 2022-09-05 chapters/background-related-work.tex
Finish document

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include the definition of centraility measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others. Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.

@@

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study \citep[see][]{cardot2011functional}.

@@

Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion. As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures. As will become apparent in the forthcoming chapters, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model. In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information. These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration. Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain (a sort of ``region selection'' algorithm).

@@

One difficulty found in almost all Bayesian methods is that the posterior distribution, the main object of interest, is usually intractable due to the integral that appears as the normalizing constant in Bayes' rule. A way to bypass this limitation is to use conjugate distributions, where it can be rigorously proven that a certain combination of likelihood and prior distributions produces a posterior distribution in the same family as the prior. However, unless conjugate priors are used, a closed-form expression is generally unattainable and some type of approximation is required. Apart from basic numerical integration, some well-performing methods in this regard are variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}. The former are based on approximating the posterior by another distribution restricted to a certain parametric family so that the Kullback-Leibler divergence between them is minimized, whereas the latter are iterative methods that directly provide approximate samples of the posterior distribution.

On a separate note, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by Bayesian neural networks. Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference approach suggested in~\citet{ma2021functional}. Moreover, there are approaches to functional regression from a Bayesian perspective that impose a Gaussian process prior on the functional parameter \(\beta\) \citep[e.g][]{lian2016posterior}, and it turns out that the RKHS's corresponding to these Gaussian processes, described for example in~\citet{van2008reproducing}, are useful when studying contraction rates of the posterior distribution.

@@

In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces. Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory. For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}. Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity. We start by defining the concept of kernel functions, which are the foundation of these spaces.

@@

Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and for simplicity assume \(\E[X(t)]=0\) for all \(t\in[0,1]\). Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\). To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional vector space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\begin{equation}\label{eq:h0}
\Hcal_0(K) = \left\{ f: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\end{equation}
Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\). However, this new space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely
\[
\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
\]
for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}: \(f(t) = \dotprod{K(t, \cdot)}{f}_K\) for all \(t \in [0, 1]\). In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.

\begin{proposition} \((\Hcal_0(K), \dotprod{\cdot}{\cdot}_K)\) is an inner product space.
\end{proposition}
  \begin{proof}
    Firstly, consider \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\), and observe that
    \[
      \dotprod{f}{g}_K = \sum_j b_j f(s_j) = \sum_i a_i g(t_i).
    \]
    These equalities show that \(\dotprod{f}{g}_K\) depends only on \(f\) and \(g\) through their values, so it is independent of their representation in \(\Hcal_0(K)\). From this expression (and also from the original definition) it is straightforward to check linearity and symmetry, while positive semidefiniteness is a direct consequence of \(K\) being a covariance function, and thus positive semidefinite. It remains to prove that \(\dotprod{f}{f}_K=0\) implies \(f=0\). Indeed, for all \(t\in[0,1]\) and \(\epsilon \in \R\), we have
    \[
    0 \leq \dotprod{f + \epsilon K(t, \cdot)}{f + \epsilon K(t, \cdot)}_K = \dotprod{f}{f}_K + 2\epsilon\dotprod{f}{K(t, \cdot)}_K + \epsilon^2 K(t, t).
    \]
    Now, if \(\dotprod{f}{f}_K=0\), by letting \(\epsilon\to 0^{\pm}\) we see that \(f(t)=\dotprod{f}{K(t, \cdot)}_K\) necessarily vanishes for all \(t\), as desired.
  \end{proof}

@@

\begin{proposition} The following properties hold for the space \(\Hcal(K)\):
  \begin{enumerate}
    \item The evaluation operator \(\delta_t\) is bounded for all \(t\in[0,1]\).
    \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
    \item If \(K\) is m-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
  \end{enumerate}
\end{proposition}
\begin{proof}
  To prove \((i)\), simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
  \[
  |\delta_t(f)| = |f(t)| = |\dotprod{f}{K(t, \cdot)}_K| \leq \|K(t, \cdot)\|_K \|f\|_K = \sqrt{K(t, t)}\|f\|_K,
  \]
  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \((ii)\), observe that for any \(t\in[0,1]\) we have
  \[
    |f_n(t) - f(t)| = |\delta_t(f_n - f)| \leq \|\delta_t\|\|f_n - f\|_K,
  \]
  as we have just shown in \((i)\) that the evaluation operator is bounded. Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform. Finally, we prove the second statement in \((iii)\), and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof. If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous. Let us now fix an arbitrary \(f\in \Hcal(K)\). By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\). But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions. Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
\end{proof}

@@

 Note that, as expected, the covariance function of \(X\) plays a crucial role in characterizing the associated RKHS. An integral operator closely related to the kernel is the so-called covariance operator, namely
\[
\K f(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
\]
which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}, so for the remaining of this work we will indeed suppose that \(K\) is continuous. As it turns out, the covariance function admits a spectral decomposition in terms of the eigenvalues and eigenfunctions of this integral operator, a sort of continuous generalization of the eigendecomposition of a symmetric positive semidefinite matrix.

@@

In connection with our RKHS theory, it can be shown that the set \(\{\sqrt{\lambda_j}\phi_j\}\) constitutes an orthonormal basis of \(\Hcal(K)\) \citep[see e.g.][Sec.~4.4]{cucker2007learning}. Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models. The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.

@@

@@

Until now we have seen that \(\Hcal(K)\) is a space related to \(X\) only through its covariance function. Nevertheless, we will now hopefully clarify that this space can indeed be seen as \textit{the} Hilbert space inherently associated with the process \(X\) itself. Let us shift the perspective momentarily and consider the linear span of \(X\) in the space of all zero-mean random variables with finite second moment, i.e.:
\[
\Lcal_0(X) = \left\{\sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
\]
It is well known that its completion in \(L^2(\Omega)\), denoted \(\Lcal(X)\), is a Hilbert space, sometimes called the Hilbert space generated by \(X\). In the words of \citet{berlinet2004reproducing}, ``\(\Lcal(X)\) \textit{contains the random variables attainable by linear operations, including limits, on the measurements of the process}''. As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\Lcal(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}. This isometry is essentially the completion of the correspondence
  \begin{equation}\label{eq:loeves-isometry}
  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
  \end{equation}
and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \Lcal(X)\). Thus, \(\Hcal(K)\) can be regarded as an isometric copy of \(\Lcal(X)\), an approach that is often useful in statistics.

However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process. It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere. An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, for each fixed \(t\) we have \(X_N(t)\to X(t)\) in the quadratic mean sense (by the Karhunen-Loève theorem), but on the other hand observe that
\[
  \E\left[\|X_N(\cdot)\|^2_K\right] = \E \left[\sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j}\right] = N \to \infty \quad (N\to\infty).
\]

@@

\begin{proposition}
  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations are satisfied:
\begin{enumerate}
  \item \(\dotprod{X}{K(t, \cdot)}_K = X(t)\), a particular reproducing property.
  \item \(\E[\dotprod{X}{f}_K] = 0\).
  \item \(\E[\dotprod{X}{f}_K\dotprod{X}{g}_K] = \dotprod{f}{g}_K\).
\end{enumerate}
\end{proposition}
Note that the statement \((iii)\) above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex feature space (e.g. infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

\begin{theorem}[Generalized representer theorem]
  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\). Then, any minimizer \(f^\ast\) of the regularized empirical risk
  \[
    R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
  \]
  is of the form \(f^\ast(\cdot) = \sum_{i=1}^n a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,n\).
\end{theorem}

@@

In the context of Bayesian inference, there are many situations in which we would like to sample from a distribution without using its explicit closed-form density, either because we do not know it, or because it is computationally difficult to do so. In these cases, we could work with a function \(f(x)\) proportional to our target density \(\pi(x)\). More specifically, in a Bayesian framework the posterior distribution is often intractable due to the normalizing integral constant, but we do know that \(posterior \propto prior \times likelihood\).

@@

>>>>

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others. Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.

@@

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and uniqueness of \(\beta\) need further study \citep[see][]{cardot2011functional}.

@@

Another set of techniques extensively studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_j)\}\) of the process that better summarize it according to some optimality criterion. As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs, bueno2019variable}, but in general they have their own dedicated algorithms and procedures. As will become apparent in the forthcoming chapters, given the nature of our suggested Bayesian model we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_j\}\), and thus provide a Bayesian-motivated variable selection process along with the other prediction methods that naturally arise within our model. In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information. These points-of-impact selection models for functional predictors have also been considered in the  literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration. Another example of a related strategy is the work of \citet{james2009functional}, in which the authors propose a method to estimate \(\beta(t)\) in such a way that it is exactly zero over some regions in the domain (a sort of ``region selection'' algorithm).

@@

One difficulty found in almost all Bayesian methods is that the posterior distribution, the main object of interest, is usually intractable due to the integral that appears as the normalizing constant in Bayes' rule. A way to bypass this limitation is to use conjugate distributions, where it can be rigorously proven that a certain combination of likelihood and prior distributions produces a posterior distribution in the same family as the prior. However, unless conjugate priors are used, a closed-form expression of the posterior is generally unattainable and some type of approximation is required. Apart from basic numerical integration, some well-performing methods in this regard are variational inference approaches \citep[e.g.][]{blei2017variational} and MCMC methods \citep[e.g.][]{brooks2011handbook}. The former are based on approximating the posterior by another distribution restricted to a certain parametric family so that the Kullback-Leibler divergence between them is minimized, whereas the latter are iterative methods that directly provide approximate samples of the posterior distribution.

On a separate note, there are some recent works that tackle Bayesian inference from a functional perspective, mostly in relation to the distribution over functions induced by a Bayesian neural network. Some examples are the functional Bayesian neural networks proposed by~\citet{sun2019functional}, variational implicit processes \citep{ma2019variational} and their deep variants \citep{ortega2022deep}, or the functional variational inference techniques suggested in~\citet{ma2021functional}. Moreover, there are approaches to functional regression from a Bayesian perspective that impose a Gaussian process prior on the functional parameter \(\beta\) \citep[e.g][]{lian2016posterior}, and it turns out that the RKHS's corresponding to these Gaussian processes, described for example in~\citet{van2008reproducing}, are useful when studying contraction rates of the posterior distribution.

@@

In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces. Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory. Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity; for a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}. We start by defining the concept of kernel functions, which are the foundation of these spaces.

@@

Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and for simplicity assume \(\E[X(t)]=0\) for all \(t\in[0,1]\). Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\). To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional vector space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\begin{equation}\label{eq:h0}
\Hcal_0(K) = \left\{ f: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\end{equation}
Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\). However, this new space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely
\[
\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
\]
for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}:
\[
f(t) = \dotprod{K(t, \cdot)}{f}_K, \quad \text{for all } t \in [0, 1].
\]
In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.

\begin{proposition} \((\Hcal_0(K), \dotprod{\cdot}{\cdot}_K)\) is an inner product space.
\end{proposition}
  \begin{proof}
    Firstly, consider \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\), and observe that
    \[
      \dotprod{f}{g}_K = \sum_j b_j f(s_j) = \sum_i a_i g(t_i).
    \]
    These equalities show that \(\dotprod{f}{g}_K\) depends only on \(f\) and \(g\) through their values, so it is independent of their representation in \(\Hcal_0(K)\). From this expression (and also from the original definition) it is straightforward to check linearity and symmetry, while positive semidefiniteness is a direct consequence of \(K\) being a covariance function, and thus positive semidefinite. It remains to prove that \(\dotprod{f}{f}_K=0\) implies \(f=0\). Indeed, for all \(t\in[0,1]\) and \(\epsilon \in \R\), we have
    \[
    0 \leq \dotprod{f + \epsilon K(t, \cdot)}{f + \epsilon K(t, \cdot)}_K = \dotprod{f}{f}_K + 2\epsilon\dotprod{f}{K(t, \cdot)}_K + \epsilon^2 K(t, t).
    \]
    Now, if \(\dotprod{f}{f}_K=0\), by letting \(\epsilon\to 0^{\pm}\) it follows that \(f(t)=\dotprod{f}{K(t, \cdot)}_K\) necessarily vanishes for all \(t\), as desired.
  \end{proof}

@@

\begin{proposition} The following properties hold for the space \(\Hcal(K)\):
  \begin{enumerate}
    \item The evaluation operator \(\delta_t\) is bounded for all \(t\in[0,1]\).
    \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
    \item If \(K\) is m-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
  \end{enumerate}
\end{proposition}
\begin{proof}
  To prove \textit{(i)}, simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
  \[
  |\delta_t(f)| = |f(t)| = |\dotprod{f}{K(t, \cdot)}_K| \leq \|K(t, \cdot)\|_K \|f\|_K = \sqrt{K(t, t)}\|f\|_K,
  \]
  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \textit{(ii)}, observe that for any \(t\in[0,1]\) we have
  \[
    |f_n(t) - f(t)| = |\delta_t(f_n - f)| \leq \|\delta_t\|\|f_n - f\|_K,
  \]
  as we have just shown in \textit{(i)} that the evaluation operator is bounded. Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform. Finally, we prove the second statement in \textit{(iii)}, and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof. If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous. Let us now fix an arbitrary \(f\in \Hcal(K)\). By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\). But then \(f_n \to f\) uniformly by \textit{(ii)}, and therefore \(f\) is continuous as the uniform limit of continuous functions. Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
\end{proof}

@@

 Note that, as expected, the covariance function of \(X\) plays a crucial role in characterizing the associated RKHS. An integral operator closely related to this covariance function is the so-called covariance operator, namely
\[
\K f(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1],
\]
which is self-adjoint and compact when \(K\) is continuous \citep[e.g.][Th.~4.6.2]{hsing2015theoretical}, so for the remaining of this work we will indeed suppose that \(K\) is continuous. As it turns out, the covariance function admits a spectral decomposition in terms of the eigenvalues and eigenfunctions of this integral operator, a sort of continuous generalization of the eigendecomposition of a symmetric positive semidefinite matrix.

@@

In connection with our RKHS theory, it can be shown that the set \(\{\sqrt{\lambda_j}\phi_j\}\) constitutes an orthonormal basis of \(\Hcal(K)\) \citep[see e.g.][Sec.~4.4]{cucker2007learning}. Furthermore, a similar orthogonal decomposition holds for the stochastic process \(X\), an expansion that lies at the base of many \(L^2\) functional regression models. The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.

@@

\enlargethispage{1\baselineskip}

@@

Until now we have seen that \(\Hcal(K)\) is a space related to \(X\) only through its covariance function. Nevertheless, we will now hopefully clarify that this space can indeed be seen as \textit{the} Hilbert space inherently associated with the process \(X\) itself. Let us shift the perspective momentarily and consider the space \(L^2(\Omega)\) of all zero-mean random variables with finite second moment, endowed with the usual norm \(\|U\|^2=\E[U]^2\). Then, the linear span of \(X\) in \(L^2(\Omega)\) is given by
\[
\Lcal_0(X) = \left\{U \in L^2(\Omega): U = \sum_{i=1}^p a_i X(t_i): p\in\N, \ a_i \in \R, \ t_i \in [0,1]\right\}.
\]
It is well known that its completion in \(L^2(\Omega)\), denoted \(\Lcal(X)\), is a Hilbert space, sometimes called the Hilbert space generated by \(X\). In the words of \citet{berlinet2004reproducing}, ``\(\Lcal(X)\) \textit{contains the random variables attainable by linear operations, including limits, on the measurements of the process}''. As it happens, via \textit{Loève's isometry} \citep{loeve1948fonctions} one can establish a congruence \(\Psi_X\) between \(\Hcal(K)\) and \(\Lcal(X)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}. This isometry is essentially the completion of the correspondence
  \begin{equation}\label{eq:loeves-isometry}
  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
  \end{equation}
and can be formally defined, in terms of its inverse, as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \Lcal(X)\). Thus, \(\Hcal(K)\) can be regarded as an isometric copy of \(\Lcal(X)\), an approach that is often useful in statistics.

However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process. It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the associated RKHS are differentiable almost everywhere. A heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, for each fixed \(t\) we have \(X_N(t)\to X(t)\) in the quadratic mean sense (by the Karhunen-Loève theorem), but on the other hand observe that
\[
  \E\left[\|X_N(\cdot)\|^2_K\right] = \E \left[\sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j}\right] = N \to \infty \quad (N\to\infty).
\]

@@

\begin{proposition}
  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations are satisfied:
\begin{enumerate}
  \item \(\dotprod{X}{K(t, \cdot)}_K = X(t)\), a particular reproducing property.
  \item \(\E[\dotprod{X}{f}_K] = 0\).
  \item \(\E[\dotprod{X}{f}_K\dotprod{X}{g}_K] = \dotprod{f}{g}_K\).
\end{enumerate}
\end{proposition}
Note that the statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property. First, a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex (e.g. infinite-dimensional) feature space, such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

\begin{theorem}[Generalized representer theorem]
  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(\mathcal R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\). Then, any minimizer \(f^\ast\) of the regularized empirical risk
  \[
    \mathcal R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
  \]
  is of the form \(f^\ast(\cdot) = \sum_{i=1}^n a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,n\).
\end{theorem}

@@

In the context of Bayesian inference, there are many situations in which we would like to sample from a distribution without using its explicit closed-form density, either because we do not know it, or because it is computationally difficult to do so. In these cases, we could work with a function \(f(x)\) proportional to our target density \(\pi(x)\). More to the point, in a Bayesian framework the posterior distribution is often intractable due to the normalizing integral constant, but we do know that \textit{posterior} \(\propto\) \textit{prior} \(\times\) \textit{likelihood}.

@@

\vspace{1em}
